Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Abstract & Intro

\begin{abstract}
Volunteer computing is used routinely
% (Paloma) Adverbs are always written before verbs, check this throughout the text
 in science and industry, but it
still has a long way to go if it is going to become
% (Paloma) I would propose "a long way to go in order to become"
 a mainstream
technology in the same way the cloud technologies in general have. The
lack of % (Paloma) Add "both" here to better complete the sentence
general purpose and open source tools with analytics built in
% (Paloma) "built-in" is an adjective and, as such, has to be written before names
could be one of the reasons why this is so, as well as the lack of
proofs of concept. That is why in this paper we present {\sf NodIO}, a
framework used to create client-server architectures for distributed
evolutionary algorithms, written entirely % (Paloma) Adverb. I would also add an "and" after the comma
in JavaScript, which needs
minimal server support to work and is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser, as well as any other client consuming its REST application
programming interface.  The architecture, along with some initial
results on the expected performance and scaling behavior when the
number of users is increased is presented. We also use
the algorithm data to find out what kind of patterns the user behavior
follow by examining the resources they devote to our experiment, and
try to fit them to some statistical distribution to create the bases
of a model of user behavior in volunteer computing experiments of this
kind that can be used by researchers to predict its performance. 
In general, our experiments prove, first, that NodIO is a flexible and
low-overhead platform to perform distributed computation experiments
that include volunteers, and then that user behavior follows some patterns,
and, thus, are amenable to a model.
\end{abstract}

@@

There is, then, space for systems that do not make any of these
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. {\sf NodIO} is first
% (Paloma) Is the "first" word really needed?
 a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can
% (Paloma) be embedded in?
 embedded any web-enabled device, including many over which 
the developers have no control.

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance costs and
extended downtime resulting from the continuous failures of existing
clusters.
Considering this, we leveraged 
% (Paloma) What 'something' did you leverage?
our experience in the design of web
applications, including JavaScript, since their inception and other
volunteer and
unconventional distributed evolutionary computing systems, and we
decided to design and release a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.
% (Paloma) I think you should rewrite this whole sentence.
% For starters, focus in that you "leveraged" the problem through your deeply knowledge of... this and that. And "Thus/Then/Therefore", you decided...

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer frameworks written in compiled languages,
such as C++ or Java, that can achieve better performance than
JavaScript. However, one of the advantages of our system is that
clients do not need to install anything in their machines, they just
need to open a web page  
% Somewhat redundant 
% Javascript is the only language that can be executed in this way without any installation, 
% also removed the default, redundant with next sentence
in a compatible browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Other advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options (such as Java Applets, Flash scripts or
ActiveX)
% (Paloma) Brackets not really needed here
. Also, as we discuss in this paper, it can provide some extra
features such as enhanced security or operating system independence.

%PABLO: hay que enlazar con el siguiente párrafo
Desktop clients have the same issues as any 
volunteer computing system, including our % (Paloma) their?
 own: these issues are
related to security as well as
safety, which introduce 
% (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
a {\em social} aspect to the design of a
computing system that makes it a {\em techno-social system}
\cite{vespignani2009predicting}. In this paper, our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page that runs the system
determine its performance and how it can be modeled and, eventually,
predicted.

The rest of the paper is organized as follows: Next we present the
state of the art (Section \ref{sec:soa}) in web-based distributed
computation systems along with attempts to predict and model its behavior. 
Then, a detailed description of the proposed system is presented
in Section \ref{sec:description}, followed by the baseline experiments
in Section \ref{sec:experiments}, and experiments using a new client
architecture that uses web workers in Section \ref{sec:w2}.
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.

>>>>

\begin{abstract}
Volunteer computing is extensively used 
% (Paloma) Adverbs are always written before verbs, check this throughout the text
in science and industry, but it
still has a long way to go in order to become
% (Paloma) I would propose "a long way to go in order to become"
 a mainstream
technology in the same way the cloud technologies in general have. The
lack of % (Paloma) Add "both" here to better complete the sentence
general purpose and open source tools with built-in analytics 
% (Paloma) "built-in" is an adjective and, as such, has to be written before names
could be one of the reasons of why this is so, as well as the lack of
proofs of concept. That is why in this paper we present {\sf NodIO}, a
framework used to create client-server architectures for distributed
evolutionary algorithms, entirely written  % (Paloma) Adverb. I would also add an "and" after the comma
in JavaScript, which needs
minimal server support to work and is able to support persistent,
asynchronous, distributed evolutionary algorithms running in the
browser, as well as any other client consuming its REST application
programming interface.  The architecture, along with some initial
results on the expected performance and scaling behavior when the
number of users is increased is presented. We also use
the algorithm data to find out what kind of patterns the user behavior
follow by examining the resources they devote to our experiment, and
try to fit them to some statistical distribution to create the bases
of a model of user behavior in volunteer computing experiments of this
kind that can be used by researchers to predict its performance. 
In general, our experiments prove, first, that NodIO is a flexible and
low-overhead platform to perform distributed computation experiments
that include volunteers, and then that user behavior follows some patterns,
and, thus, are amenable to a model.
\end{abstract}

@@

There is, then, space for systems that do not make any of these
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can
% (Paloma) be embedded in?
be embedded in any web-enabled device, including many over which 
the developers have no control.

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance costs and
extended downtime resulting from the continuous failures of existing
clusters.
Considering this, we leveraged 
% (Paloma) What 'something' did you leverage?
our experience in the design of web
applications, including JavaScript, since their inception and other
volunteer and
unconventional distributed evolutionary computing systems to 
design and release a new free framework that would allow anyone to
create a volunteer distributed EC experiment using cloud resources as
servers and browsers as clients.
% (Paloma) I think you should rewrite this whole sentence.
% For starters, focus in that you "leveraged" the problem through your deeply knowledge of... this and that. And "Thus/Then/Therefore", you decided... 
% (Mario) We leverage our experience in blah blah to design

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer frameworks written in compiled languages,
such as C++ or Java, that can achieve better performance than
JavaScript. However, one of the advantages of our system is that
clients do not need to install anything in their machines, they just
need to open a web page  
% Somewhat redundant 
% Javascript is the only language that can be executed in this way without any installation, 
% also removed the default, redundant with next sentence
in a compatible browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Other advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX.
% (Paloma) Brackets not really needed here
Also, as we discuss in this paper, it can provide some extra
features such as enhanced security or operating system independence.

%PABLO: hay que enlazar con el siguiente párrafo
%Desktop clients have the same issues as any 
%volunteer computing system, including our % (Paloma) their?
% own: these issues are
%related to security as well as
%safety, which introduce

When developing volunteer based systems, desingers must
also consider the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. 
The computing system becomes a {\em techno-social system} \cite{vespignani2009predicting}.  
% (Paloma) Is the social aspect related to the word "safety"? It's not completely clear.
%a {\em social} aspect to the design of a
%computing system that makes it a {\em techno-social system}
In this paper, our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page running the system
determine its performance and how it can be modeled and, eventually,
predicted.

The rest of the paper is organized as follows: Next we present the
state of the art (Section \ref{sec:soa}) in web-based distributed
computational systems along with attempts to predict and model its behavior. 
Then, a detailed description of the proposed system is presented
in Section \ref{sec:description}, followed by the baseline experiments
in Section \ref{sec:experiments}, and experiments using a new client
architecture that uses web workers in Section \ref{sec:w2}.
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.
